---
title: "Introduction to airHealth"
author: "İbrahim Oğuz Asan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to airHealth}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `airHealth` package provides tools for analyzing air pollution data and its health impacts. This vignette demonstrates the main features and functionality of the package.

## Installation

```{r install, eval = FALSE}
# Install from GitHub
devtools::install_github("yourusername/airHealth")
```

## Air Quality Analysis

### Calculating Air Quality Index (AQI)

The package provides functions to calculate AQI based on multiple pollutants:

```{r aqi_example}
library(airHealth)

# Example data
pm25 <- c(12, 35, 55, 150, 250)
pm10 <- c(54, 154, 254, 354, 424)
o3 <- c(54, 70, 85, 105, 200)
no2 <- c(53, 100, 360, 649, 1249)

# Calculate AQI
aqi_results <- calculate_aqi(pm25, pm10, o3, no2)
print(aqi_results)
```

### Health Impact Assessment

We can estimate health impacts based on AQI and population data:

```{r health_impact}
# Example data
population <- 100000
exposure_hours <- 24

# Calculate health impact
health_impact <- calculate_health_impact(aqi_results$AQI)
print(health_impact)
```

### Respiratory Health Effects

The package can estimate respiratory health effects from air pollution:

```{r respiratory}
# Estimate respiratory effects
population <- 100000
respiratory_effects <- estimate_respiratory_effects(pm25, pm10, population)
print(respiratory_effects)
```

## Visualization

### AQI Trends

```{r aqi_plot}
# Create example time series data
aqi_data <- data.frame(
  date = seq.Date(from = as.Date("2024-01-01"), by = "day", length.out = 10),
  AQI = c(50, 100, 150, 200, 300, 250, 200, 150, 100, 50)
)

# Plot AQI trend
plot_aqi_trend(aqi_data, "date")
```

### Health Impact Visualization

```{r health_plot}
# Create example health impact data
health_data <- data.frame(
  Category = c("Low", "Moderate", "High", "Very High", "Extreme"),
  Health_Impact_Score = c(500, 2500, 7500, 25000, 75000)
)

# Plot health impact
plot_health_impact(health_data)
```

### Correlation Analysis

```{r correlation}
# Create example correlation data
set.seed(123)
air_quality_data <- data.frame(
  PM2.5 = rnorm(100),
  PM10 = rnorm(100),
  O3 = rnorm(100),
  NO2 = rnorm(100),
  Health_Score = rnorm(100)
)

# Plot correlation heatmap
plot_correlation_heatmap(air_quality_data)
```

## Best Practices

When using the package, consider the following best practices:

1. Always validate your input data before analysis
2. Use appropriate units for pollutant concentrations
3. Consider local air quality standards and guidelines
4. Account for population demographics in health impact assessments
5. Regularly update your data sources

## References

1. US EPA Air Quality Index (AQI) Basics
2. World Health Organization Air Quality Guidelines
3. Global Burden of Disease Study 